---
title: "How to Use the commonsMath Package"
author: "David B. Dahl"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{How to Use the commonsMath Package}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
scalaAvailable <- ! isTRUE(tryCatch({rscala::scalaConfig(reconfig="offline")}, error=function(e) TRUE))
knitr::opts_chunk$set(echo = TRUE, eval=scalaAvailable)
```

## Overview

The commonsMath package provides access to the [Apache Commons Mathematics Library](http://commons.apache.org/proper/commons-math/).  It can
can accessed via the:

+ [rscala](https://cran.r-project.org/package=rscala) package, or
+ [rJava](https://cran.r-project.org/package=rJava) package.

We demonstrate below how to use it in:

+ R scripts, or
+ R packages

## Usage in R Scripts

#### Access via rscala
```{r}
library("rscala")
s <- scala("commonsMath")
rng1 <- s$.new_org.apache.commons.math3.random.RandomDataGenerator()
rng1$reSeed(7342L)
rng1$nextGaussian(0,1)
```

#### Access via rJava

```{r}
library("rJava")
.jinit(Sys.glob(file.path(system.file(package="commonsMath"), "java", "*.jar")))
rng2 <- .jnew("org.apache.commons.math3.random.RandomDataGenerator")
rng2$reSeed(.jlong(7342L))
rng2$nextGaussian(0,1)
```

## Usage in R Packages

#### Access via rscala

The `DESCRIPTION` should have `Imports: rscala, commonsMath`.

The `NAMESPACE` should have `import(rscala)`.

Define an `.onLoad` function like the following:

```{r}
.onLoad <- function(libname, pkgname) {
  s <- scala("commonsMath")
  assign("s", s, envir = parent.env(environment()))
}
```

Package functions can then assess classes and methods from the commonsMath package, e.g.:

```{r}
rstdnorm <- function() {
  rng1 <- s$.new_org.apache.commons.math3.random.RandomDataGenerator()
  rng1$nextGaussian(0.0,1.0)
}
```

#### Access via rJava

The `DESCRIPTION` should have `Imports: rJava, commonsMath`.

The `NAMESPACE` should first have `import(rscala)` and then have `import(commonsMath)`.

Define an `.onLoad` function like the following:

```{r}
.onLoad <- function(libname, pkgname) {
  .jpackage(pkgname, lib.loc=libname)
}
```

Package functions can then assess classes and methods from the commonsMath package, e.g.:

```{r}
rstdnorm <- function() {
  rng2 <- .jnew("org.apache.commons.math3.random.RandomDataGenerator")
  rng2$nextGaussian(0,1)
}
```




